Cultural Transmission and Technological Transitions during the Late Paleolithic in Korea

### 2022-10-15 analysis/paper/paper.qmd
restore missing plot

Compared to Phase 2, the influence of tang length and stem length of Phase 3 becomes more pronounced and while stem width is less effective (@fig-pca-attributes). However, there are no notable changes in correlations over time.

We explored assemblages that contain multiple stemmed points to examine the temporal patterns with the premise that stemmed points from one assemblage were made by the same group of people. There are four assemblages, three assemblages from the Suyanggae site (SYG) and one from the Yongsandong site (YS) that contain seven or more stemmed points. SYG1_2 and SYG6_2 are from Phase 2 and SYG1_3 and YS_3 are from Phase 3. Following the previous studies of [@bettinger1997evolutionary; @bettinger1999point; @eerkens2008cultural; @mesoudi2008cultural; @garvey2018current], we only compared CV values of attributes, body length, maximum length, mid width, and tang width. @fig-eight shows that CV values for the four assemblages are mostly below 25, which represent transmissions through indirect bias. Among the assemblages, SYG1_2 (n=7) has higher CV values as well as the widest range of confidence intervals. Compared to the Phase 2, CV values of Phase 3 are lower, which is different from @fig-plot-cv-over-time results of body length and maximum length from all assemblages. It is assumed that those attributes became more standardized within the same assemblages. Compared to Suyanggae assemblages, Yongsandong has smaller CV values.

![CV values of four attributes from assemblages that have 7 or more stemmed points. Among the four assemblages, SYG1_2 and SYG6_2 are from Phase 2 of the Korean Late Paleolithic chronology and SYG1_3 and YS_3 are from Phase 3.](images/006-cv-four-assemblage.png){#fig-eight width="386"}

@@

>>>>

We then explored assemblages that contain multiple stemmed points to examine the temporal patterns with the premise that stemmed points from one assemblage were made by the same group of people. There are four assemblages, three assemblages from the Suyanggae site (SYG) and one from the Yongsandong site (YS) that contain more than 5 stemmed points. SYG1_2 and SYG6_2 are from Phase 2 and SYG1_3 and YS_3 are from Phase 3. Following the previous studies of [@bettinger1997evolutionary; @bettinger1999point; @eerkens2008cultural; @mesoudi2008cultural; @garvey2018current], we only compared CV values of attributes, body length, maximum length, mid width, and tang width. @fig-plot-cv-four-assemblage shows that CV values for the four assemblages are mostly below 25, which represent transmissions through indirect bias. Among the assemblages, SYG1_2 (n=7) has higher CV values as well as the widest range of confidence intervals. Compared to the Phase 2, CV values of Phase 3 are lower, which is different from @fig-plot-cv-over-time results of body length and maximum length from all assemblages. It is assumed that those attributes became more standardized within the same assemblages. Compared to Suyanggae assemblages, Yongsandong has smaller CV values.

```{r}
#| label: fig-plot-cv-four-assemblage
#| fig-cap: CV values of four attributes from assemblages that more than 5 stemmed points. Among the four assemblages, SYG1_2 and SYG6_2 are from Phase 2 of the Korean Late Paleolithic chronology and SYG1_3 and YS_3 are from Phase 3.
#| fig-width:   4.5
#| fig-height:  4.5

#### plot CVs by site with number of artefacts showing >5

cv_by_site_df_label_more_than_5 <-
  df_sitename %>%
  unite(col = "site_phase",
        c(sitename,
          SPstage.Stage),
        sep = "_") %>%
  dplyr::select(-SPstage.Raw_material) %>%
  group_by(site_phase) %>%
  add_tally() %>%
  filter( n > 5) %>%
  mutate(label = as.factor(paste0(site_phase,' (N = ', n, ")"))) %>%
  ungroup() %>%
  dplyr::select(-n, -site_phase) %>%
  pivot_longer(-label,
               names_to = "variable",
               values_to = "value") %>%
  group_by(label, variable) %>%
  summarise(cv_by_site =   sharma_cv(value),
            cv_low_sharma =  sharma_int_low(value),
            cv_high_sharma = sharma_int_high(value)) %>%
  filter(variable %in% c("ML", "BL","MW", "TW"))

cv_by_site_df_label_more_than_5 %>%
  ggplot(aes(x = factor(variable),
             y = cv_by_site)) +
  geom_point() +
  geom_errorbar(aes(ymin = cv_low_sharma,
                    ymax = cv_high_sharma), width = 0.2) +
  xlab("Attribute") +
  ylab("Coefficient of Variation on Attributes") +
  facet_wrap( ~ label, scales = "free_y") +
  coord_cartesian(ylim = c(15,60)) +
  theme_bw()

@@

Compared to Phase 2, the influence of tang length and stem length of Phase 3 becomes more pronounced and while stem width is less effective. However, there are no notable changes in correlations over time.
